---
title: Warm Up 1
subtitle: Functions in Shiny Apps
output: 
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE}
library(shiny)
```


When is the last time you thought about using functions in Shiny apps? Or about Shiny apps in general? Complete these four thought exercises for a quick warm up on the topic.

# Thought Exercise 1

The code below creates a Shiny app. Can you tell what the app will do? Will the app function correctly when you run the code?

```{r eval = FALSE}
ui <- fluidPage(
  sliderInput("slider", "Slide Me", 0, 100, 1),
  textOutput("num")
)

server <- function(input, output) {
  output$num <- renderText({
    input$slider
  })
}

shinyApp(ui, server)
```

You bet it will! The app contains a simple slider, and displays the value of the slider in a reactive way. Now let's change things up a bit.

# Thought Exercise 2

This app (below) attempts to do the same thing as the app above, but it does it in a different way. Can you spot the change? It tries to use a function to create the slider input. 

Will the app work correctly when you run the code? (If you are playing along at home, don't forget to stop the app above before running the app below).

```{r eval = FALSE}
slider <- function() {
  sliderInput("slider", "Slide Me", 0, 100, 1)
}

ui <- fluidPage(
  slider(),
  textOutput("num")
)

server <- function(input, output) {
  output$num <- renderText({input$slider})
}

shinyApp(ui, server)
```

If you said "Yes, that app will work," you are right. You can write functions that return html output for Shiny, as `slider()` does, and then place the output straight into your UI. 

If you would like a function to return multiple elements for the UI, wrap the elements in `tagList()`, e.g.

```{r eval = FALSE}
elements <- function() {
  tagList(
    sliderInput("slider", "Slide Me", 0, 100, 1),
    numericInput("number", "Number Me", 0)
  )
}
```

# Thought Exercise 3

Now let's be ambitious, and add two sliders (and two outputs) to our app. Will this app work correctly when you run the code?

```{r eval = FALSE}
slider <- function() {
  sliderInput("slider", "Slide Me", 0, 100, 1)
}

ui <- fluidPage(
  slider(),
  textOutput("num1"),
  slider(),
  textOutput("num2")
)

server <- function(input, output) {
  output$num1 <- renderText({input$slider})
  output$num2 <- renderText({input$slider})
}

shinyApp(ui, server)
```

No. The app is messed up. Don't believe me? Move the first slider. *Both* numbers change. Move the second slider. *Nothing* changes. 

What has gone wrong? Each input object in a Shiny app, like a slider, needs a unique id. But our `slider()` function returns a slider with the id `"slider"` *every time you call it*. As a result, Shiny cannot track the values in a reasonable way.

Surely there is a way to fix the app. Can you think of one?

# Thought Exercise 4

Here's one way (to try). Will this app work correctly when you run the code?

```{r eval = FALSE}
slider <- function(id) {
  sliderInput(id, "Slide Me", 0, 100, 1)
}

ui <- fluidPage(
  slider("slider1"),
  textOutput("num1"),
  slider("slider2"),
  textOutput("num2")
)

server <- function(input, output) {
  output$num1 <- renderText({input$slider1})
  output$num2 <- renderText({input$slider2})
}

shinyApp(ui, server)
```

Yes, it will. Here we arrange for `slider()` to create a slider with a unique id each time we call it.

As a bonus, we see that you can pass and use arguments in functions that create Shiny output. And why wouldn't you? That's what R functions are all about.

We're now ready to tacke Shiny Modules. See you back at the [article](Big-apps.html).

